Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

### 2023-04-23 Draft.Rmd
pred test train

The Extreme Gradient Boosting (XGB) model performs by training weak learners, models that have low prediction accuracy, sequentially to become a single strong learner, a model that has strong prediction accuracy. We preprocess the training set to make the predictive modeling process more accurate and smoother by prepping (estimates the quantities for 'train_scaled') and baking (assigns prepped data to an object) the data. The aforementioned top 10 predictors were used in this preprocessing recipe. Moreover, the categorical variables were factored and features with no variance were removed before this data was baked. Then, this processed data is assigned to 'folds3' and resampled on 5 subsets of the training set. Using 'fold3', we tune the hyperparameters to make the modeling process more efficient and perform grid specification to determine which hyperparameter values have high prediction accuracy (i.e. low prediction error). Next, we isolated the best hyperparameter values, so they can be used for the final boosting model. Lastly, we fitted the model and evaluated its prediction performance. Its RMSE score was 0.56, indicating the XG Boosting is a good model. This website was used: [RBloggers](https://www.r-bloggers.com/2020/05/using-xgboost-with-tidymodels/)

@@

### 4. Random FOrest Regression

@@

```{r}
# Create table of prediction metrics from training data
linreg = res1 |> collect_metrics() |> as.data.frame()
linreg = cbind(linreg[,1], round(linreg[,3],2))
knn = res2 %>% collect_metrics() |> as.data.frame()
knn = cbind(knn[,1], round(knn[,3],2))
xgboost = xgboost_score_train |> filter(.metric!='mae') |> as.data.frame()
xgboost = cbind(xgboost[,1], xgboost[,3])
ranfor = rf_score_train |> filter(.metric!='mae') |> as.data.frame()
ranfor = cbind(ranfor[,1], ranfor[,3])

@@

@@

@@

As seen on the table, LR has the highest RMSE and lowest r-squared estimate. KNN has the second highest RMSE with a moderate R-squared estimate. And XGB has the lowest RMSE of 0.56 and the highest R-squared estimate of 0.96. Random Forest model has the both the second lowest RMSE and second highest R-squared estimate. Since we are choosing the best model by the best RMSE, which is the lowest one, the Xgboost model is thus the best and final model based on our procedure. After fitting the Xgboost model to the testing data, we find that its RMSE and R-squared is 0.61 and 0.94, respectively. Since we are trying to predict annual average ambient air pollution of a given monitor, then these statistics mean that the amount of error in our model's predictions is about 0.61 $\mu$g/m$^3$, and that the features we used in our model are able to explain 94% of the variance in the data, which means this model fits/predicts the data well.

## Results
```{r Results}
compare |> ggplot(aes(y=rmse)) + geom_bar()

@@

```{r Q3}

@@

Given that the top 10 features were 'CMAQ,' 'lat,' 'county_area,' 'lon,' 'aod,' 'popdens_county,' 'zcta,' 'log_nei_2008_pm25_sum_10000,' 'fips,' and 'log_nei_2008_pm10_sum_10000,' we think that XG Boost model will perform well since the original dataset included states of varying sizes with different climates and pollution levels. Moreover, its RMSE score is 0.62 and R-squared is 0.94, indicating that the prediction performance of this model is good in relation to the other models. Since Hawaii and Alaska are almost polar opposites with the former being geographically smaller and having hotter climate, we suspect that the model will be able to predict PM2.5 concentration levels in Hawaii and Alaska if they were included in this dataset since the diversity but also inter-relatedness of the features in this dataset aids the model in predicting ambient air pollution in differing environments.

>>>>

The Extreme Gradient Boosting (XGB) model performs by training weak learners, models that have low prediction accuracy, sequentially to become a single strong learner, a model that has strong prediction accuracy. We preprocess the training set to make the predictive modeling process more accurate and smoother by prepping (estimates the quantities for 'train_scaled') and baking (assigns prepped data to an object) the data. The aforementioned top 10 predictors were used in this preprocessing recipe. Moreover, the categorical variables were factored and features with no variance were removed before this data was baked. Then, this processed data is assigned to 'folds3' and resampled on 5 subsets of the training set. Using 'fold3', we tune the hyperparameters to make the modeling process more efficient and perform grid specification to determine which hyperparameter values have high prediction accuracy (i.e. low prediction error). Next, we isolated the best hyperparameter values, so they can be used for the final boosting model. Lastly, we fitted the model and evaluated its prediction performance. Its RMSE score was 0.55, indicating the XG Boosting is a good model. This website was used: [RBloggers](https://www.r-bloggers.com/2020/05/using-xgboost-with-tidymodels/)

@@

### 4. Random Forest Regression

@@

```{r Best&Final}
# Create table of prediction metrics from training data
linreg = res1 |> collect_metrics() |> as.data.frame()
linreg = cbind(linreg[,1], round(linreg[,3],2))
knn = res2 %>% collect_metrics() |> as.data.frame()
knn = cbind(knn[,1], round(knn[,3],2))
xgboost = xgboost_score_train |> filter(.metric!='mae') |> as.data.frame()
xgboost = cbind(xgboost[,1], xgboost[,3])
ranfor = rf_score_train |> filter(.metric!='mae') |> as.data.frame()
ranfor = cbind(ranfor[,1], ranfor[,3])

@@

# Visual of rmse for our 4 models
compare |> ggplot(aes(x=model, y=rmse)) + 
  geom_bar(stat="identity") + 
  labs(title='RMSE of the Four Models') +
  theme_minimal()

@@

# Get pred. metrics based on testing data
test_prediction = xgboost_model_final |> 
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000, data=train_processed) |> 
  predict(new_data=test_processed) |> 
  bind_cols(testing(dat_split))

@@

As seen on the table and visual, LR has the highest RMSE and lowest r-squared estimate. KNN has the second highest RMSE with a moderate R-squared estimate. And XGB has the lowest RMSE of 0.55 and the highest R-squared estimate of 0.96. Random Forest model has the both the second lowest RMSE and second highest R-squared estimate. Since we are choosing the best model by the best RMSE, which is the lowest one, the Xgboost model is thus the best and final model based on our procedure. After fitting the Xgboost model to the testing data, we find that its RMSE and R-squared is 1.23 and 0.73, respectively. Since we are trying to predict annual average ambient air pollution of a given monitor, then these statistics mean that the amount of error in our model's predictions is about 1.23 $\mu$g/m$^3$, and that the features we used in our model are able to explain 73% of the variance in the data, which means this model fits/predicts the data well.

@@

```{r Q3}
# Fit RF model to testing data

xgb_score_test = test_prediction |> 
  yardstick::metrics(value, .pred) |> 
  mutate(.estimate=format(round(.estimate, 2), big.mark=","))
xgb_score_test

@@

Given that the top 10 features were 'CMAQ,' 'lat,' 'county_area,' 'lon,' 'aod,' 'popdens_county,' 'zcta,' 'log_nei_2008_pm25_sum_10000,' 'fips,' and 'log_nei_2008_pm10_sum_10000,' we think that XG Boost model will perform well since the original dataset included states of varying sizes with different climates and pollution levels. Moreover, its RMSE score is 0.42 and R-squared is 0.98, indicating that the prediction performance of this model is excellent. Since Hawaii and Alaska are almost polar opposites with the former being geographically smaller and having hotter climate, we suspect that the model will be able to predict PM2.5 concentration levels in Hawaii and Alaska if they were included in this dataset since the diversity but also inter-relatedness of the features in this dataset aids the model in predicting ambient air pollution in differing environments.
